Adopting agile in the large: motivations, success factors and challenges

### 2014-05-07 journal/article.tex
Work in progress

The difficulty of agile and misunderstandings were also evident in cases where
the methods were customized badly. As by-the-book implementation often was not
feasible the agile method was tailored to suite the needs of the organization.
However, in some cases tailoring meant skipping some practices, which lead into
problems. In one case certain individuals were allowed to ignore some core
elements of Scrum, which came to turn the teams decision making into a variant
of command and control [P9].
Lewis and Neher [P29] describe how there was temptation to strip some practices
and enhance others. Previous attempts had proven that one of the reasons early
attempts of agile had failed was deviations from the process, because of which
the agile mindset did not take root [P29]. In a bad implementation customization
may lead to the team taking only practices that reflect their current needs, and
therefore failing to achieve a real change in process and mindset [P9, P46].

@@

@@

* Hard to create roadmap between teams --> Teams see only their own priorities

@@

@@

* Coordination challenge --> Team goal or product goal?
* Team centric focus created strong boundaries --> Bad for coordination
[P7]
A challenge we faced in team coordination was understanding whether to optimize
the team goal or the product goal. Some teams focused too much on their own
goals and created strong boundaries, impeding coordination.

* ?
[P10]
CROSS-FUNCTIONAL DEPENDENCIES: However well we did within the team, we also
frequently ran into dependencies on other groups often in other locations –
groups such as UED (User Experience and Design), operations, legal and the
Yahoo! security team (affectionately called the Paranoids) – which were
identified as blocking issues but almost never neatly resolved within our sprint
boundaries.

* Well into the transformation there was trouble with dependencies between teams
* Management burden to manage dependencies --> Creates problem with roadmapping
* Hard to create roadmap between teams --> Teams see only their own priorities
* Reduce dependencies --> Help roadmapping
[P10]
Now  teams were  comprised of tight-knit groups that – while collocated wherever
possible – were  responsible to deliver  value regardless of location,
technology or historical allegiance.
However, within six months or so, and despite a pattern  of  success and
improvement, it became very clear that the new forking between horizontal
and vertical  teams  placed an enormous burden on the product and program
management teams to manage dependencies in a way that supported  the  broader
organizational roadmap.  Teams were structured to execute independently,  but
the broader product roadmap called for features that  implied  specific platform
improvements. Establishing a  reliable roadmap therefore required massive (and 
often unreliable) coordination between teams that were more often  than not
head-down in their own (sub-sprint) priorities.
Theoretically,  the product owners (assisted by the Music PMO) would ensure that
capabilities required by the  roadmap would appear on their own team’s backlog
at an appropriate level of priority.  In practice, it was very difficult to
identify  and track such dependencies, especially when the  feature  set  (and
associated set of dependencies) was constantly in flux.
Phase 2 of the transition (continuing today –  see Figure 4) therefore  attempts
to further reduce such dependencies and improve how they are managed over time
to improve release and roadmap planning.

* Team centricity --> Power --> However, they operate with independent priorities --> No respect for larger context (release)
* Release Framework was created
[P13]
Shortly after the initial rollout to 30+ product teams it was clear that we had
unleashed the power of our individual teams through their new found
self-organization and focus on short sprints.  However, since we aggregate the
work of many sprints into a single release it was also becoming clear that our
scrum teams were operating independently without respect for the context of the
larger release.  Herding 30+ teams working in the same codeline to a predictable
time-boxed release date is very challenging in any traditional development model
let alone an agile development model where we value self-organization over
centralized control.
Our answer to this problem was a lightweight Release Framework and the
introduction of a Release Manager for the overall product release.

@@

* Scrum on team level --> Software interfaces not properly synchronized
* Problems with larger features
* Scrum on team level good --> True benefit requires entire program to be flexible for changes
[P26]
Using scrum only on the team level had led to additional problems. Software
interfaces were not properly synchronized with other development teams.
Larger features were causing synchronization problems due to dependencies in
code. Synchronization to the main waterfall program was considered problematic.
Deploying scrum on the team level had brought flexibility in the team, but the
ability to embrace the change was not truly benefiting the whole program as it
was lacking capabilities to systematically embed the changes in all teams and
levels.

* Different customers compete --> Customer centricity does not view the system as a whole
[P29]
each customer base views the system only via their process flow.
We have multiple competing customers crossing multiple co-dependent
applications, including dependencies on applications outside of our
organization; multiple funding sources; and a mixture of Waterfall and Scrum
projects.
For our software teams to be able to execute efficiently, we will need to push
the complexity and prioritization upstream, scaling teams to meet demand while
maintaining clean, simple prioritized Product Backlogs for each product.

@@

[P43] Ryan 2008 - The Price of Agile Is Eternal Vigilance.pdf - 43:11 [We implemented a scrum-of-scru..]  (4:48-4:200)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
No memos

We 
implemented a scrum-of-scrums run by our release 
management team to identify inter-team dependencies 
and resolve merge issues before they occur.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:18 [What could not be mitigated we..]  (4:2133-4:2292)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

What could not be mitigated were issues with teams dependent on “unreliable”
Scrum teams, at least if the Scrum teams were isolated in a non-Scrum environment.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:28 [Several approaches are known t..]  (8:522-8:1054)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [P: Scrum of scrums - Family: Practices] 
No memos

Several approaches are known to co-ordinate the work of several Scrum teams working
on one product, e.. Scrum of Scrums [9] or MetaScrum [18]. The size and complexity
of SAP’s products makes it necessary to consider these techniques. To co-ordinate the
work of multiple Scrum teams and to communicate the requirements via team product
backlogs, various approaches were combined. Special product teams were formed for
this purpose. These teams can be regarded as both a permanent Scrum of Scrums and a
supervisory product owner.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:32 [Product teams were introduced ..]  (8:1076-8:1627)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Product teams were introduced as a second organizational layer above the Scrum teams.
A product team is responsible for the work of up to 7 development teams. It consists
of the product owners of those teams and the same number of team members who
are specialists in particular ﬁelds. Depending on the problem area other experts may
be included. This allows for full engineering coverage of all problem areas expected,
well-organized communication among the teams, and direct communication with the
development teams to detect and mitigate risks.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:33 [Product teams can rarely suppo..]  (9:841-9:1526)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Product teams can rarely support more than seven Scrum teams. Ifthe products or components are be too big or too complex to be developed in this organizational setup, an
intermediate organizational layer is inserted between the product team and the Scrum
teams. These teams are called “area product teams”. Their responsibility is the same as
the product teams’ responsibility with the exception of budget and ﬁnal product decisions. In this three-layer organization a productteam works similar to a project management ofﬁce. The complete budget and product responsibility is with the product team,
while the operational day-to-day decisions are handled by the area product teams.

[P45] Schnitter 2011 - Large-Scale Agile Software Development At Sap Ag.pdf - 45:38 [Scaling Scrum to a multi-team ..]  (10:1223-10:1784)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] [O: Large scale definition - Family: Contextual] 
No memos

Scaling Scrum to a multi-team development organization is not as easy. Scrum of
Scrums and backlog preparation for many teams can be combined by product teams
as outlined above. Looking at all known parameters of agile project management at
SAP we tend to say that not much more than 130 development employees may be
organizedin that fashion. This number sums up developers in 7 teams (max. 70 people),
the product team (max. 16), developmentinfrastructure responsibles (about 10), quality
assurance and testers (about 25), general management (about 10).

[P48] Smits 2011 - Agile Experience Report - Transition And Complexity At Cisco Voice Technology Group.pdf - 48:30 [Tracking progress of hundreds ..]  (4:1657-4:2111)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Tracking progress of hundreds of teams working on 
pieces of major features proved to be a major 
problem. Tools that we were using did not have roll 
up capability without imposing strict rules on how 
data was entered, labeled and updated. We found out 
our tools were not really scalable for the size and 
complexity of the programs we were working on. 
This proved a serious handicap in providing 
executive summary information about progress.

[P50] Sutherland 2011 - Hitting The Wall - What To Do When High Performing Scrum Teams Overwhelm Operations And Infrastructure.pdf - 50:20 [A key impediment was the distr..]  (4:1737-4:2643)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

A key impediment was the distributed 
development model in use by the Scrum teams.  
Originally, before Scrum adoption, all teams worked
together in a common environment.  There were 
many integration issues as a result and no solid 
process to rectify them.  So with the initial adoption 
of Scrum in summer 2009 it was decided that each 
individual Scrum team should have a dedicated and 
private server with integration to take place after the 
build.  In retrospect in the fall of 2009, the downside 
of the decision to break up the environment was 
being felt as integration issues were being discovered 
very late,  as  the  build  process  was  the  first  point  of 
integration.  To overcome this impediment the 
decision was made to re-integrate all teams onto a 
single  server  environment  again,  so  that  integration 
issues could be rectified before they left 
development.

[P51] Vlaanderen 2012 - Growing Into Agility - Process Implementation Paths For Scrum.pdf - 51:12 [Initially, the sprints lasted ..]  (5:230-5:717)   (Super)
Codes:  [C: Coordinating/integrating work of many teams - Family: RQ3: Challenges] 
No memos

Initially, the sprints lasted two or three weeks, depending on
the team, causing a discrepancy between the sprint durations of some of the
teams. At the same time, interdependencies started to become more of an issue
as the amount of teams grew. This caused problems when, for instance, frontend teams needed functionality in the back-end that was not yet developed.
These problems, which consisted mainly of delays, were solved by ﬁxating sprint
duration at two weeks for all teams.

@@

>>>>

The difficulty of agile and misunderstandings were also evident in cases where
the methods were customized badly. As by-the-book implementation often was not
feasible the agile method was tailored to suit the needs of the organization.
However, in some cases tailoring meant skipping some practices, which lead into
problems. In one case certain individuals were allowed to ignore some core
elements of Scrum, which came to turn the teams decision making into a variant
of command and control [P9].
Lewis and Neher [P29] describe how there was temptation to strip some practices
and enhance others. Previous attempts had proven that one of the reasons early
attempts of agile had failed was deviations from the process, because of which
the agile mindset did not take root [P29]. In a bad implementation customization
may lead to the team taking only practices that reflect their current needs, and
therefore failing to achieve a real change in process and mindset [P9, P46].

@@

% The software produced in a large organization may have size and complexity that
% agile struggles to handle.

The primary studies revealed challenges on two areas which relate particularly
to large scale. First, a large software development endeavour must handle
integrating the work products of many teams. Second, a large organization will
have management challenges not seen on small scale.

@@

% Hard to create roadmap between teams --> Teams see only their own priorities

Some cases reported that introducing agile practices soon yielded improvements
on the team level. However, the initial success was followed by challenges when
the teams needed to work as parts of the larger surrounding orgaization [P13,
P17, P24]. Introducing agile had created flexibility on a team level, but the
flexibility could not be fully utilized as the surrounding organization was not
responsive enough [P26]. Even being agile, teams had dependencies to each other
and the dependencies were causing problems [].

%% ?? -->
Large organization put agile methods under the test of scaling up, and problems
were encountered. One case reported use of agile across hundreds of teams, but
management could not get a roll up of the progress without having strict rules
on how status was reported [P48]. Schnitter and Mackert [P45] report that a
maximum workable size in a project managed by Scrum was 130 people including 7
development teams.

Some organizations attempted to initially create a model where teams operate
autonomously, which is well in line with agile principles. However, a number of
issues arose from independence. For instance, teams had to balance between the
team's own goals and broader goals of the entire organization, and some teams
chose focus on their own goal [P7]. Coordination was obstructed by independent
teams that did not respect the larger context they were working in [P13].
Coordination was also difficult when different teams were working for different
customers but the applications being built were interdependent [P29]. One case
reported that allowing teams to have different sprint durations created delays
in delivery [P51].

One case reported that team centricity created also distrust between teams [P24].

@@

DEPENDENCIES

* Dependencies --> Because of a general roadmap
* Dependencies --> Teams must work for a general higher level goal
* Management burden to manage dependencies --> Creates problem with roadmapping
* Hard to create roadmap between teams --> Teams see only their own priorities
* Reduce dependencies --> Help roadmapping
[P10]
Now teams were comprised of tight-knit groups that – while collocated wherever
possible – were responsible to deliver value regardless of location, technology
or historical allegiance.
However, within six months or so, and despite a pattern of success and
improvement, it became very clear that the new forking between horizontal and
vertical teams placed an enormous burden on the product and program management
teams to manage dependencies in a way that supported the broader organizational
roadmap. Teams were structured to execute independently, but the broader product
roadmap called for features that implied specific platform improvements.
Establishing a reliable roadmap therefore required massive (and often
unreliable) coordination between teams that were more often than not head-down
in their own (sub-sprint) priorities.
Theoretically, the product owners (assisted by the Music PMO) would ensure that
capabilities required by the roadmap would appear on their own team’s backlog at
an appropriate level of priority. In practice, it was very difficult to identify
and track such dependencies, especially when the feature set (and associated set
of dependencies) was constantly in flux.
Phase 2 of the transition (continuing today – see Figure 4) therefore attempts
to further reduce such dependencies and improve how they are managed over time
to improve release and roadmap planning.

CODE SYNCHRONIZATION

* Added problems by agile --> Software interfaces not properly synchronized
* Problems with larger features --> dependencies in code
* Synchronization to main waterfall problematic
[P26]
Using scrum only on the team level had led to additional problems. Software
interfaces were not properly synchronized with other development teams.
Larger features were causing synchronization problems due to dependencies in
code. Synchronization to the main waterfall program was considered problematic.
Deploying scrum on the team level had brought flexibility in the team, but the
ability to embrace the change was not truly benefiting the whole program as it
was lacking capabilities to systematically embed the changes in all teams and
levels.

* Problems with inter-team coordination --> Architecture fragile, code dialects, distrust between teams
* (Multisite, teams on different sites)
[P24]
No matter how much communication we tried to establish between the teams, we
were still far away from shared team focus on the complete product. As a result
we could see the architecture becoming more fragile, multiple code dialects and
distrust between teams were also starting to grow.
We still needed to improve the focus on the complete product and not on "bad
code” from other teams. The main barrier for building quality into the product
was the distrust between the teams who worked on the same product. Therefore the
two R\&D teams in Denmark and Bangladesh were merged into one global team.
Instead of two distributed teams, working with different sprint backlogs, we
established one team with one sprint backlog. The problem of building quality
into the complete product was now a shared team problem, not a management
problem.

* With Scrum each team started using their own integration server --> Problem with late integration
* The shared integration server was taken back into use
[P50]
A key impediment was the distributed development model in use by the Scrum
teams. Originally, before Scrum adoption, all teams worked together in a common
environment. There were many integration issues as a result and no solid process
to rectify them. So with the initial adoption of Scrum in summer 2009 it was
decided that each individual Scrum team should have a dedicated and private
server with integration to take place after the build. In retrospect in the fall
of 2009, the downside of the decision to break up the environment was being felt
as integration issues were being discovered very late, as  the build process was
the first point of integration. To overcome this impediment the decision was
made to re-integrate all teams onto a single server environment again, so that
integration issues could be rectified before they left development.

OTHER ``DEPENDENCIES''

COORDINATION

* Managers --> Getting big picture, overview
* Coordination challenges across teams
* Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
[P17]
As individual component teams began to experience success, the challenge changed
to coordinating across teams. Although iterations were humming along at the team
level it was initially difficult for managers to get a complete picture of how
the release was progressing. They were unable to tell when particular features
were complete as their requirements crossed team boundaries.
Using Rally’s application for managing the development lifecycle, BMC is now
able to guide the overall release objectives with themes and high-level
features.

* Size and complexity --> Scrum-of-Scrums is mandatory 
[P45]
The size and complexity of SAP’s products makes it necessary to consider these
techniques. To co-ordinate the work of multiple Scrum teams and to communicate
the requirements via team product backlogs, various approaches were combined.
Special product teams were formed for this purpose. These teams can be regarded
as both a permanent Scrum of Scrums and a supervisory product owner.

* Problem: Tracking the work of hundreds of teams --> Handicap in providing executive summary on progress
[P48]
Tracking progress of hundreds of teams working on pieces of major features
proved to be a major problem. Tools that we were using did not have roll up
capability without imposing strict rules on how data was entered, labeled and
updated. We found out our tools were not really scalable for the size and
complexity of the programs we were working on. This proved a serious handicap in
providing executive summary information about progress.

@@

HELP / SOLUTIONS

* Release Framework was created
[P13]
Shortly after the initial rollout to 30+ product teams it was clear that we had
unleashed the power of our individual teams through their new found
self-organization and focus on short sprints.  However, since we aggregate the
work of many sprints into a single release it was also becoming clear that our
scrum teams were operating independently without respect for the context of the
larger release.  Herding 30+ teams working in the same codeline to a predictable
time-boxed release date is very challenging in any traditional development model
let alone an agile development model where we value self-organization over
centralized control.
Our answer to this problem was a lightweight Release Framework and the
introduction of a Release Manager for the overall product release.

@@

* Scrum of scrums to resolve inter-team dependencies
[P43]
We implemented a scrum-of-scrums run by our release management team to identify
inter-team dependencies and resolve merge issues before they occur.

* Solution: Product teams
[P45]
Product teams were introduced as a second organizational layer above the Scrum
teams. A product team is responsible for the work of up to 7 development teams.
It consists of the product owners of those teams and the same number of team
members who are specialists in particular fields. Depending on the problem area
other experts may be included. This allows for full engineering coverage of all
problem areas expected, well-organized communication among the teams, and direct
communication with the development teams to detect and mitigate risks.

% --

@@

% INDIVIDUAL TEAMS GOOD --> PROBLEMS WHEN MANY TEAMS NEED TO WORK TOGETHER
% 
% * Individual component teams gained success --> Coordination challenges across teams
% * Managers have difficult to get complete picture of the progress of the release --> Features crossing team boundaries
% * Help: Tool support to visualize dependencies for people
% [P17]
% As individual component teams began to experience success, the challenge changed
% to coordinating across teams. Although iterations were humming along at the team
% level it was initially difficult for managers to get a complete picture of how
% the release was progressing. They were unable to tell when particular features
% were complete as their requirements crossed team boundaries.
% Using Rally’s application for managing the development lifecycle, BMC is now
% able to guide the overall release objectives with themes and high-level
% features.
% 
% 
% * Team centricity --> Power --> However, they operate with independent priorities --> No respect for larger context (release)
% * Release Framework was created
% [P13]
% Shortly after the initial rollout to 30+ product teams it was clear that we had
% unleashed the power of our individual teams through their new found
% self-organization and focus on short sprints.  However, since we aggregate the
% work of many sprints into a single release it was also becoming clear that our
% scrum teams were operating independently without respect for the context of the
% larger release.  Herding 30+ teams working in the same codeline to a predictable
% time-boxed release date is very challenging in any traditional development model
% let alone an agile development model where we value self-organization over
% centralized control.
% Our answer to this problem was a lightweight Release Framework and the
% introduction of a Release Manager for the overall product release.
% 
% 
% * Initial adopting good
% * Problems with inter-team coordination --> Architecture fragile, code dialects, distrust between teams
% * (Multisite, teams on different sites)
% * Help: Retrospectives across teams --> Created global teams --> Shared backlog
% [P24]
% To embrace and implement Scrum across the different teams and organizations, we
% focused on doing regular retrospectives on both team and company level.
% The first couple of company retrospectives were quite hectic and required some
% extra facilitation techniques to be constructive. But we did identify some of
% the main problems when crossing the team boundaries and working together with
% other departments.
% No matter how much communication we tried to establish between the teams, we
% were still far away from shared team focus on the complete product. As a result
% we could see the architecture becoming more fragile, multiple code dialects and
% distrust between teams were also starting to grow.
% By doing regular retrospectives with all the different teams and across the
% company we identified the main problems and decided it was time to create the
% global Scrum teams. We wanted to improve communication and trust by working on
% shared commitments between the different teams.
% 
% We still needed to improve the focus on the complete product and not on "bad
% code” from other teams. The main barrier for building quality into the product
% was the distrust between the teams who worked on the same product. Therefore the
% two R\&D teams in Denmark and Bangladesh were merged into one global team.
% Instead of two distributed teams, working with different sprint backlogs, we
% established one team with one sprint backlog. The problem of building quality
% into the complete product was now a shared team problem, not a management
% problem.
% 
% 
% * Scrum on team level --> Software interfaces not properly synchronized
% * Problems with larger features
% * Scrum on team level good --> True benefit requires entire program to be flexible for changes
% [P26]
% Using scrum only on the team level had led to additional problems. Software
% interfaces were not properly synchronized with other development teams.
% Larger features were causing synchronization problems due to dependencies in
% code. Synchronization to the main waterfall program was considered problematic.
% Deploying scrum on the team level had brought flexibility in the team, but the
% ability to embrace the change was not truly benefiting the whole program as it
% was lacking capabilities to systematically embed the changes in all teams and
% levels.
%

@@

* Transformation of an organization to use agile can be seen as transforming
  also the agile method to suit the organization.
  --> A large scale agile model may require tailoring and trial over time

* It can be argued that any development effort with hundreds of teams will be
  difficult to coordinate

* Even though most of the primary studies are experience reports and subject to
  bias, this study can draw valuable conclusions. --> If a matter is being
  talked about it must be of some significance.
